### 2022-06-09 inst/jma_rebuttal_review3.tex
newlines to mark end of my blue comments

\documentclass[12pt]{article}
\usepackage{xcolor}
\begin{document}

(Below, the reviewer's comments are in black, and my replies to the
issues are in \textcolor{blue}{blue}.  I have indicated changes to the
manuscript where appropriate.

\section*{Rebuttal to second third review: overall response}

Short story: the reviewers make reasonable and constructive comments.
I have accommodated the comments with  with rewording.  The resulting
document is, I believe, stronger and more scholarly than before and I
recommend it to you.

\section*{Detailed rebuttal: Reviewer \#1}

Comments from the Editors and Reviewers:

Thank you for this revised manuscript.  My apologies for contributing
to the slow turnaround of these reviews.

I have given this version the official status of ``Minor Revision" to
reflect the fact that we all agree that this draft is much closer to
publication-ready. In particular, I believe that the remaining
reviewer concerns can be addressed with moderate rewriting.  However,
some of the reviewer feedback falls closer to ``Major Revision," so I
wanted to acknowledge that a few of the remaining critiques are beyond
surface level.

As you will see in the detailed feedback below, one of the reviewers
still has substantive concerns about the overall framing of your
discussion and how directly it addresses the mathematical and
aesthetic choices that went into the software package.  I suspect that
you can address the concerns in the first few review paragraphs by
being more specific about the constraints of your implementation, the
level of flexibility you hoped to provide for the user, and the
alternate possibilities for future work, but since I don't know what
technical decisions went into your coding, I might be misjudging
exactly what is suitable.

I look forward to receiving your further revision, and will make sure
the editorial process goes more quickly from this point forward.

Best wishes,

Susan Goldstine, Ph.D.
Associate Editor
Journal of Mathematics and the Arts

Reviewer \#1: The paper has improved in terms of its overall
presentation in the context of knot renderings, and many detailed
fixes have been made as suggested by the reviewers.

However, the paper still has the major weakness of being too much a
somewhat ad-hoc technical manual on ``How-to-use-R and Inkplot to get
some decent knot plots" by using the software package [27].

The paper is lacking good discussions of the key issues and
explanations of the choices made, specifically: What makes a knot plot
attractive, followed by a thorough discussion of how to best set up a
knot representation and an optimization procedure to achieve the
desired qualities?

\textcolor{blue}{I appreciate the sentiment, which I take to be a
  comment on the large amount of programming-oriented detail in the
  submission.  This reviewer evidently regards the knots in the
  submission as ``decent'', which I take to be a great compliment.  It
  is often difficult to know the correct balance between overly
  technical descriptions [which indeed often overlaps with the aims of
    a technical manual] and gushing conceptual big-picture thinking.
  However, I note that many of the comments below are essentially
  requests for clarifications of minor technical details of the
  implementation: and this would indicate that perhaps the balance
  should be shifted back toward the technical manual end of the
  spectrum.  The other observation I would make here is that the
  software package [27] is intended to be part of the submission (it
  is released under the GPL and is thus available for public use) and
  with this understanding, even a downbeat technical manual would have
  scholarly value in conjunction with the new software it describes.
  I want the editor and reviewer to know that in addressing the
  concerns below I have tried hard to answer not just ``what'' but
  also the ``why'' of the implementation.  \\ \\ Nevertheless, the cue
  above is a good one.  The two questions: (a) {\em what makes a knot
    plot attractive?}, and (b) {\em how best ... to achieve this?}
  are indeed profound, and were not addressed directly in the old
  manuscript.  In the revision I have prominently stated the questions
  in the introduction, and given a provocative, if brief, outline of
  my view.\\ \\}

The problem starts with the basic curve representation:

Why use Bezier curves?

\textcolor{blue}{Most of the answer is in the manuscript: ``A Bezier
  curve is a visually pleasing polynomial path that can be used to
  specify the path of a knot projection; here cubic Bezier curves are
  used''.  However, Bezier curves are a natural choice for a number of
  reasons and I have added the following sentence: ``Bezier curves are
  a natural choice for working with mathematical knots for several
  reasons: they are familiar to many graphical workers; they have a
  natural and intuitive control system (usually called ``handles''),
  and are implemented in many graphical software suites''.  I do not
  wish to suggest that Bezier curves are the only parametrizations for
  working with knots; but they are a natural starting point, and
  AFAICS they have no serious disadvantages in this context, nor is
  there any obvious superior method.\\ \\}

Why not use some other spline curve that guarantees G2-continuity?

\textcolor{blue}{I don't have a good answer to that, other than to
  point out that the focus of the manuscript was to investigate the
  visual appearance of knots using mathematical optimization routines.
  I do not wish to rule out using other G2-continuous splines, and it
  might be interesting some day to consider (e.g.) B-splines.  But as
  I say above, Bezier curves are a good starting point.\\ \\}

And, if forced by the chosen tool to use Bezier curves, why not place
nodes at the crossing points, (and one or two more for large loopy
lobes between subsequent crossing points)?  -- This would allow the
user to place the crossing points far enough apart to start with, and
this would make it easy to set the Bezier handles to force right-angle
crossings.

\textcolor{blue}{Most of this issue is discussed below, at the comment
  for page 4, lines 53-54.  But here I would observe that it is not at
  all easy to manipulate Bezier handles on two distinct nodes [which
    happen to coincide spatially] to result in their two respective
  lines being orthogonal.  Observe that the overstrand and understrand
  of a crossing point may have (for good reasons!) very different
  curvatures.\\ \\}

Also, the implementation in the ``badness function" seems far from
optimal.  ``Total crossing angles()" does not address a key problem
directly: If one wants to avoid highly acute crossing angles, then
this can be addressed in a much more sensitive way with a function
such as 1/ sqr( cross.angle ).  This would go to infinity in the worst
case!  It also corresponds more directly to the way that excessive
curvature is avoided by minimizing ``bending energy," which is the
integral over the square of curvature -- or: 1/ sqr( curve-radius).
These are some ``math" issues that readers of JMA would be interested
in.

\textcolor{blue}{Function {\tt total\_crossing\_angles()} was poorly
  named.  With crossing angles $\theta_i$, $i=1,2,\ldots k$ the
  function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
  \\ \\}

Some more detailed comments:

Page 3 line 42: ``...knots possess a line of symmetry (at least, the
diagrams do if the breaks are ignored), which ..."  -$>>$ ``...knots
possess a C2 rotation axis, which ..."

Page 4 lines 22-25: ``One plausible technique for creating knot
projections is to consider a two-dimensional projection of a knot's
embedding E in R3. However, this approach often results in poor visual
appearance: cusps or other displeasing effects can occur."  -$>>$ True!
- But what is the alternative, if the user has an entangled physical
knot in her hands?

\textcolor{blue}{Being able to render a nice optimized diagram
  corresponding to a physical knot that one has in one's hands is a
  reasonable wish.  But not one that I can fulfil.  Indeed, the
  package is predicated upon the notion that the diagram is the
  ``real'' object, and the (3D) knot itself is irrelevant.  Thus much
  of the package {\em prevents} the optimization routine from making
  Reidermeister moves [that is, changes that affect the topology of
    the diagram but not that of the 3D knot].  The operational reason
  for this is that the user specifies a particular diagram which the
  software then optimizes for appearance.  The manuscript includes two
  diagrams of $10_{125}$, the Perko pair, which show that sometimes
  one wants to preserve the topology of the diagram.\\ \\}

Page 4 line 45: Define ``nodes" !

\textcolor{blue}{``Nodes'' is a standard word in the context of Bezier
  curves. I have not used the word in any other sense in the
  manuscript and I think the text is clear as it stands.\\ \\}

Page 4 line 48: ``However, if we can ensure that the radius of
curvature changes only by a small amount,..."  -$>>$ HOW can the user do
this?

\textcolor{blue}{It is not the user that stops large changes, it is
  the optimization routine and in particular the rather eclectic set
  of badness functions.  Incidentally, the ``keep nodes far from
  crossing points'' badness function is instrumental in stopping large
  changes of this type from occurring.\\ \\ OLD TEXT: However, if we
  can ensure that the radius of curvature changes only by a small
  amount, visual continuity---in the sense of absence of visible
  interruptions to a smoothly evolving strand---is
  preserved. \\ \\ NEW TEXT: However, because the optimization routine
  ensures that the radius of curvature changes only by a small
  amount, visual continuity---in the sense of absence of visible
  interruptions to a smoothly evolving strand---is
  preserved.\\ \\
}

Page 4 lines 50-52: No need to differentiate here between over- and
under-strands.  I assume they are treated the same in the optimization
process.

\textcolor{blue}{The referee is correct to say that the over and under
  strand are treated the same by the optimization routine, but my
  intent was to introduce and define the words themselves (rather than
  to differentiate between them).\\ \\}

Page 4 lines 53-54: ``We would like strand crossing points to be far
from nodes,..."  -- Why?  The curves are also visually continuous across
the nodes!

\textcolor{blue}{The curves certainly are visually continuous across
  the nodes---but only in the optimized knots.  The motivation for
  keeping Bezier nodes far from crossing points is twofold.  Firstly,
  as one traces a curve visually, Bezier nodes cannot possibly avoid
  introducing {\em some} form of disjointness in the path; and such
  disjointness is particularly disruptive if on the understrand of a
  node, as the visual continuity is interrupted.  I have added a
  figure (including two different diagrams of a trefoil knot) to the
  Gallery that illustrates the reason why nodes' being at crossing
  points is undesirable.  In short, at the node of a Bezier curve, the
  radius of curvature undergoes a visually jarring discontinuity which
  looks bad, especially the understrand, which is particularly
  susceptible to such considerations because of the break.  I do
  appreciate that many viewers will find the distinction invisible or
  inconsequential; but my informal polling of family and colleagues
  reveals that, once sensitised to the issue, they cannot unsee the
  kink.  \\ \\ The second reason is more mundane.  With reference to
  figure 8 (the four figure-of-eight knots diagram), we can see a
  vertical line of symmetry on which there are two crossing points.
  Imagine there was a node at one of these crossing points.  There
  would have to be {\em two} nodes at the same point, one for the
  overstrand and one for the understrand; the software would somehow
  have to impose vertical mirror symmetry on the diagram and this
  would necessitate that the knot class contain yet another symmetry
  object, alongside the 12 existing ones.  These considerations become
  yet more onerous when considering rotational symmetry as in Figure 9
  (which shows $5_1$).  If there were a node at, say, the intersection
  of strand 18-19 and 9-10 then somehow the two nodes would have to
  embody not only mirror symmetry (with respect to 5-6/14-15) but also
  fivefold rotational symmetry.  Simply put, this is a nightmare.  The
  software documentation contains a substantial discussion of the
  issue, but I do not believe that the manuscript would benefit from
  what amounts to little more than me grumbling about implementation
  difficulties.\\ \\}

Page 5 lines 46-51: Why are there SEVEN significant decimal digits?
Does the graphic screen have 10 million pixels per line?

\textcolor{blue}{This is the default print method used by R (which
  internally uses IEC-60559 arithmetic, 14 significant figures).  Any
  attempt by me to change the print method would, I believe, be
  confusing and unnecessary.  The point of the table was to show that
  a knot path could be represented by a few dozen floating-point
  numbers.\\ \\}

Figure 3, 5, 7: Perhaps fewer than 100 circles per Bezier curve would
be better.

Figure 4.  Knot $7_6$ with strands numbered... -- The main use here seems
to be to see very clearly the extent of every Bezier curve.
Over-/under-crossings seem to become relevant only for the final
rendering of the knot curve with the gaps included.

\textcolor{blue}{Good point, I have added a sentence to this effect to
  the caption of figure 4.\\ \\}

Page 7 line 52: ``The optimization typically proceeds over R 50" - What
is the meaning of this?

\textcolor{blue}{It is a 50-dimensional optimization, optimizing a
  function with fifty parameters (which correspond to different
  positions of Bezier nodes and handles).\\ \\}

Page 7 line 56: ``...function total crossing angles()" The function is
not explicitly defined in the paper, but it sounds clearly suboptimal.
One bad acute-angle crossing can easily ```hide" in a sum of all angles.
It would be better to individually penalize acute angles.  A suitable
term might be: 1/ sqr( cross.angle ).  This would go to infinity in
the worst case!

\textcolor{blue}{Function {\tt total\_crossing\_angles()} was poorly
  named.  With crossing angles $\theta_i$, $i=1,2,\ldots k$ the
  function actually returns $\sum_{i=1}^k\left(\cos\theta_i\right)^2$.
  \\ \\}

Page 8 line 20, and footnote: The ``devil" and the ``interesting math"
is in the details!  The footnote is not sufficient to explain the
trade-offs that are being made in the optimization process.

Page 9, line 62: What is ``Bosch-type symmetry:..." ?  Please explain.

\textcolor{blue}{Bosch-type symmetry refers to Bosch (2010) which was
  cited in the previous sentence.\\ \\ old text: The package
  implements Bosch-type symmetry\ldots \\ \\ new text: The package
  implements symmetry in a similar way to that of Bosch
  (2010)\ldots\\ \\}

$>>$ If all nodes fulfill the specified symmetry constraints (as shown
on page 11, lines 15-22 and on lines 35-43), but the connection
diagram does not completely adhere to the specified symmetry - What
happens? - Does the package check this and let the user know about the
discrepancy?

\textcolor{blue}{This situation cannot occur: symmetric nodes imply a
  symmetric diagram.\\ \\}

Page 10 lines 56-58: ``Minimizing the badness is not entirely
straightforward..."  -- What is the problem?  I think that if all the
points are moved jointly within the specified symmetry constraints, an
optimized symmetrical solution will result.

\textcolor{blue}{The problem is reconstructing the knot from the
  minimal set of of nodes and the various types of imposed symmetry.
  This is not difficult exactly, but not entirely straightforward
  either.\\ \\ Also, optimizing high-dimensional functions is rarely
  straightforward: the search space is large; in my field of
  computational ecology we talk of the ``curse of dimensionality''.
  One has to contend with a wide variety of frustrating issues such as
  local extrema, expensive function evaluation, and not least the
  inadequacies of Euclidean distance as a metric for the domain.
  Dimensionally cursed problems such as that addressed by {\tt knotR}
  are sometimes literally intractable.  My considered view is that
  Nelder-Mead simplex technique used in the package is unreasonably
  effective.\\ \\ }

{The above two comments suggest, that it might be advantageous to use
  some hierarchy in the representation of symmetrical knots: Only
  specify the curve for one unique fragment of the knot curve, and
  compose the complete curve by instancing that fragment with
  mirroring and rotation as required to construct the complete knot!
  The optimization would then only have to be done for this unique
  fragment (with proper end-conditions).  If this same hierarchy is
  also used in the graphical display of the knot curve, the user could
  easily make adjustments to just a few nodes, while always
  automatically obtaining the desired symmetrical result.}

\textcolor{blue}{This is exactly what I have done!  The referee gives
  an accurate summary of the code's methodology.\\ \\}

Page 10 lines 55: ``In the package, the badness weightings may be
altered easily,..."  -- What is the user interface for this?  Are
there sliders in the display that allow interactive adjustments of the
weights?

\textcolor{blue}{I have not implemented sliders.  The weightings may
  be changed in function {\tt badness()}, and documentation (including
  examples) is given at {\tt badness.Rd}.\\ \\}

Figure 9: The fact that ``angle-crossing penalty" had to be enhanced by
a factor of 100 to achieve a more desirable (in my opinion) result,
illustrates the fact that the "total crossing angles()"-function is
not a good choice for optimization.

\textcolor{blue}{As discussed above, {\tt total\_crossing\_angles()}
  was a somewhat misleading name.  The factor of 100 ensured that even
  tiny deviations from exactly $90^\circ$ were heavily penalised.\\ \\}

I am missing a description how the two breaks at every crossing point
are being graphically implemented.

\textcolor{blue}{New figure added (Figure 6) which illustrates the
  implementation\\ \\}

Conclusions:

RE: ``Further work might include functionality to deal
with links."  $>>$ Readers may think that this would be a trivial
extension.  Please add a paragraph like the one you wrote in response
to Reviewer 1, to explain why the extension to links may be more
difficult than one might think.

\textcolor{blue}{Done, new subsection ``some reflections on links''
  at the end, under ``conclusions and further work''\\ \\}

RE: ``In the broader context of optimization in art, we observe that
numerical optimization techniques can produce aesthetically pleasing
results, an observation that might find uptake by graphic artists."
$>>$ This only holds with the proviso that we can think of a good
mathematical formulations to capture the various features that capture
aesthetics appropriately.  - This then becomes the real ``art."

\textcolor{blue}{ An insightful observation.\\ \\ OLD TEXT In the
  broader context of optimisation in art, we observe that numerical
  optimisation techniques can produce aesthetically pleasing results,
  an observation that might find uptake by graphic artists.\\ \\ NEW
  TEXT In the broader context of optimisation in art, we observe that
  numerical optimisation techniques can produce aesthetically pleasing
  results, an observation that might find uptake by graphic artists.
  Indeed, one could argue that real art lies in the capturing of
  mathematical formulations of aesthetics.  \\ \\ }

Gallery:

Knot diagrams $8_3$ through $8_6$ in Figure 12 stand out with their helices
too tightly wound.  The two criteria on page 4, lines 54 and 55 do not
seem to be sufficiently enforced.  I suggest adjusting these figures
in the spirit of Figure 9b.  This may happen automatically with a
better ``badness function" that penalizes non-orthogonal crossing
angles more directly.

\textcolor{blue}{The referee is correct to say that changing the
  badness function would result in figures less like 9a and more like
  9b.  But I would take issue with the phrase ``too tightly wound''.
  I have polled my colleagues who appear to be evenly split between
  preferring 9a and 9b.  The comparison appears to elicit strong
  opinions here at AUT.  Proponents of 9b point to the broader winding
  along the central helix, while those of 9a point to the more liquid
  and languid appearance, and in particular emphasise the absence of
  retrograde curvature at the bottom.  I have added a brief discussion
  of these issues to the caption of figure 12 [which is now figure 15
    of the resubmission].  I would observe that the {\tt knotR}
  package makes such discussions possible.\\ \\}

[27]:  Incomplete reference.  Perhaps give a URL.

\textcolor{blue}{Canonical URL now added\\ \\}

Reviewer \#2: Firstly I would like to apologise for my late response on
this. Reviewing this edited paper had not got onto my radar.

Secondly all the points in my original report have now been
addressed. I do feel that the literature review could have gone a
little bit further beyond my stated suggestions, but it is now
sufficient to give the paper more context.

\end{document}